---
title: "Build Demographics from Table 1"
author: "Thomas Castro"
date: "11/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(kableExtra)

set.seed(2021)

```

## Data

Data from **Virtual care expansion in the Veterans Health
Administration during the COVID-19 pandemic: clinical
services and patient characteristics associated with
utilization** Table 1.


### Veterans
Veterans in DC: 31252

National Averages for Use of Virtual Care
22.4% - Never 
62.5% - Existing 
15.1% - New 

```{r}

veterans_num <- 31252
VID <- sample(10000:90000, veterans_num )

# Use category
use <- c( rep("Never", 224), rep("Existing", 625), rep("New", 151) )
use <- sample(use, veterans_num, replace = TRUE)

never <- sum(use == "Never")
existing <- sum(use == "Existing")
new <- sum(use == "New")

```


### Age Group
```{r}
# Within Never
subgroup <- c( rep("18-44", 19), rep("45-64", 26), rep("65 plus", 55))
age <- sample(subgroup, never, replace = TRUE)

# Within Existing
subgroup <- c( rep("18-44", 18), rep("45-64", 31), rep("65 plus", 52))
age <- c(age, sample(subgroup, existing, replace = TRUE) )

# Within New
subgroup <- c( rep("18-44", 18), rep("45-64", 30), rep("65 plus", 53))
age <- c(age, sample(subgroup, new, replace = TRUE) )
  
```



### Race
```{r}
# Based on https://planning.dc.gov/sites/default/files/dc/sites/op/page_content/attachments/DC%20Veterans%20Statistics%20%28003%29.pdf

# Within Never - ESTIMATES FOR CODING PURPOSES
subgroup <- c( rep("White", 421), rep("Black", 552), rep("Asian", 30),
               rep("Hispanic", 48), rep("Other", 17))
race <- sample(subgroup, never, replace = TRUE)

# Within Existing - ESTIMATES FOR CODING PURPOSES
subgroup <- c( rep("White", 321), rep("Black", 552), rep("Asian", 30),
               rep("Hispanic", 48), rep("Other", 17))
race <- c(race, sample(subgroup, existing, replace = TRUE) )

# Within New - ESTIMATES FOR CODING PURPOSES
subgroup <- c( rep("White", 521), rep("Black", 552), rep("Asian", 30),
               rep("Hispanic", 48), rep("Other", 17))
race <- c(race, sample(subgroup, new, replace = TRUE) )
  
```


### Sex
```{r}
# Within Never
subgroup <- c( rep("Male", 934), rep("Female", 66))
sex <- sample(subgroup, never, replace = TRUE)

# Within Existing
subgroup <- c( rep("Male", 900), rep("Female", 100))
sex <- c(sex, sample(subgroup, existing, replace = TRUE) )

# Within New
subgroup <- c( rep("Male", 919), rep("Female", 81))
sex <- c(sex, sample(subgroup, new, replace = TRUE) )
  
```


### Marital Status
```{r}
# Within Never
subgroup <- c( rep("Single", 380), rep("Married", 600), rep("Unknown", 18))
marital <- sample(subgroup, never, replace = TRUE)

# Within Existing
subgroup <- c( rep("Single", 460), rep("Married", 530), rep("Unknown", 13))
marital <- c(marital, sample(subgroup, existing, replace = TRUE) )

# Within New
subgroup <- c( rep("Single", 400), rep("Married", 590), rep("Unknown", 15))
marital <- c(marital, sample(subgroup, new, replace = TRUE) )
  
```


### Priority Group
```{r}
# Within Never
subgroup <- c( rep("No Service Disability", 23), rep("Low Income", 16), 
               rep("Low/Moderate Disability", 31),rep("High Disability", 30))
priority <- sample(subgroup, never, replace = TRUE)

# Within Existing
subgroup <- c( rep("No Service Disability", 14), rep("Low Income", 19), 
               rep("Low/Moderate Disability", 22),rep("High Disability", 45))
priority <- c(priority, sample(subgroup, existing, replace = TRUE) )

# Within New
subgroup <- c( rep("No Service Disability", 17), rep("Low Income", 16), 
               rep("Low/Moderate Disability", 25),rep("High Disability", 41))
priority <- c(priority, sample(subgroup, new, replace = TRUE) )
  
```


### Homeless condition
.98% of Never
5.7% of Existing
1.4% of New
```{r}
# Within Never
subgroup <- c( rep(TRUE, 98), rep(FALSE, 9902))
homeless <- sample(subgroup, never, replace = TRUE)

# Within Existing
subgroup <- c( rep(TRUE, 57), rep(FALSE, 943))
homeless <- c(homeless, sample(subgroup, existing, replace = TRUE) )

# Within New
subgroup <- c( rep(TRUE, 14), rep(FALSE, 986))
homeless <- c(homeless, sample(subgroup, new, replace = TRUE) )
  
```


### Chronic conditions
1-2 conditions 50% of all veterans
3-4 conditions 31% of all veterans
5+  conditions 19% of all veterans
```{r}
# Within 1 - 2 (guessed at 65/35 split)
subgroup <- c( rep(1, 65), rep(2, 35))
chronic <- sample(subgroup, never, replace = TRUE)

# Within 3 - 4 (guessed at 50/50 split)
subgroup <- c( rep(3, 50), rep(4, 50))
chronic <- c(chronic, sample(subgroup, existing, replace = TRUE) )

# Within 5 + (guessed)
subgroup <- c( rep(5, 50), rep(6, 35), rep(7, 20), rep(8, 10), rep(9, 5))
chronic <- c(chronic, sample(subgroup, new, replace = TRUE) )

```


### Mental Health condition
17% of Never
41% of Existing
32% of New
```{r}
# Within Never
subgroup <- c( rep(TRUE, 17), rep(FALSE, 83))
mental <- sample(subgroup, never, replace = TRUE)

# Within Existing
subgroup <- c( rep(TRUE, 41), rep(FALSE, 59))
mental <- c(mental, sample(subgroup, existing, replace = TRUE) )

# Within New
subgroup <- c( rep(TRUE, 32), rep(FALSE, 68))
mental <- c(mental, sample(subgroup, new, replace = TRUE) )

```


```{r}

veterans_table1 <- tibble(VID = VID,
                          use = use,
                          age = age,
                          race = race,
                          sex = sex,
                          marital = marital,
                          priority = priority,
                          homeless = homeless,
                          chronic = chronic,
                          mental = mental)

head(veterans_table1)

```

```{r}

write_csv(veterans_table1, "csv_files/veterans_table1.csv")

```

