---
title: "Sample Veteran Data Demographics Summary Report"
author: "Telehealth Team"
date: "11/4/2021"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(kableExtra)

# colorblind friendly palette
cp <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
        "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

demo <- read_csv("csv_files/demographics.csv")

demo <- demo %>% 
  arrange(use_of_care) %>% 
  mutate(use_of_care = factor(use_of_care, levels=c("virtual_never",
                                            "virtual_existing",
                                            "virtual_new")))
```
Summary of the demographic information/distribution of the sample DC Veteran dataset.  All demographic designations are based on the nationwide averages reported in table 1 of **Virtual care expansion in the Veterans Health Administration during the COVID-19 pandemic: clinical services and patient characteristics associated with utilization**


$~$

$~$

$~$

$~$

## Age
```{r, message=FALSE}

demo %>% 
  group_by(use_of_care, age) %>% 
  summarise(group_n = n()) %>%
  mutate(percentage = sprintf("%0.1f%%", (group_n/sum(group_n) * 100))) %>% 

  kable(align=rep('c', 4)) %>% 
  kable_styling(bootstrap_options = "striped") %>% 
  pack_rows("Virtual - Never", 1, 3,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - Existing", 4, 6,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - New", 7, 9,
            label_row_css = "background-color: #666; color: #fff;")

demo %>% 
  group_by(use_of_care) %>% 
  count(age) %>% 
  ggplot(aes(x = use_of_care,
             y = n,
             fill = age) ) +
  geom_bar(position = "fill",
           stat = "identity") +
  coord_flip() + scale_fill_manual(values=cp) +
  scale_y_continuous(labels = scales::percent_format()) 
  
```

$~$

$~$

$~$

$~$

## Sex
```{r, message=FALSE}

demo %>% 
  group_by(use_of_care, sex) %>% 
  summarise(group_n = n()) %>%
  mutate(percentage = sprintf("%0.1f%%", (group_n/sum(group_n) * 100))) %>% 
  kable(align=rep('c', 4)) %>% 
  kable_styling(bootstrap_options = "striped") %>% 
  pack_rows("Virtual - Never", 1, 2,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - Existing", 3, 4,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - New", 5, 6,
            label_row_css = "background-color: #666; color: #fff;")

demo %>% 
  group_by(use_of_care) %>% 
  count(sex) %>% 
  ggplot(aes(x = use_of_care,
             y = n,
             fill = sex) ) +
  geom_bar(position = "fill",
           stat = "identity") +
  coord_flip() + scale_fill_manual(values=cp) +
  scale_y_continuous(labels = scales::percent_format())

```

$~$

$~$

$~$

$~$

## Race
```{r, message=FALSE}

demo %>% 
  group_by(use_of_care, race) %>% 
  summarise(group_n = n()) %>%
  mutate(percentage = sprintf("%0.1f%%", (group_n/sum(group_n) * 100))) %>% 
  kable(align=rep('c', 4)) %>% 
  kable_styling(bootstrap_options = "striped") %>% 
  pack_rows("Virtual - Never", 1, 6,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - Existing", 7, 12,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - New", 13, 18,
            label_row_css = "background-color: #666; color: #fff;")

demo %>% 
  group_by(use_of_care) %>% 
  count(race) %>% 
  ggplot(aes(x = use_of_care,
             y = n,
             fill = race) ) +
  geom_bar(position = "fill",
           stat = "identity") +
  coord_flip() + scale_fill_manual(values=cp) +
  scale_y_continuous(labels = scales::percent_format())

```

$~$

$~$

$~$

$~$

## Ethnicity
```{r, message=FALSE}

demo %>% 
  group_by(use_of_care, ethnicity) %>% 
  summarise(group_n = n()) %>%
  mutate(percentage = sprintf("%0.1f%%", (group_n/sum(group_n) * 100))) %>% 
  kable(align=rep('c', 4)) %>% 
  kable_styling(bootstrap_options = "striped") %>% 
  pack_rows("Virtual - Never", 1, 3,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - Existing", 4, 6,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - New", 7, 9,
            label_row_css = "background-color: #666; color: #fff;")

demo %>% 
  group_by(use_of_care) %>% 
  count(ethnicity) %>% 
  ggplot(aes(x = use_of_care,
             y = n,
             fill = ethnicity) ) +
  geom_bar(position = "fill",
           stat = "identity") +
  coord_flip() + scale_fill_manual(values=cp) +
  scale_y_continuous(labels = scales::percent_format())

```

$~$

$~$

$~$

$~$

## Marital Status
```{r, message=FALSE}

demo %>% 
  group_by(use_of_care, marital) %>% 
  summarise(group_n = n()) %>%
  mutate(percentage = sprintf("%0.1f%%", (group_n/sum(group_n) * 100))) %>% 
  kable(align=rep('c', 4)) %>% 
  kable_styling(bootstrap_options = "striped") %>% 
  pack_rows("Virtual - Never", 1, 3,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - Existing", 4, 6,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - New", 7, 9,
            label_row_css = "background-color: #666; color: #fff;")

demo %>% 
  group_by(use_of_care) %>% 
  count(marital) %>% 
  ggplot(aes(x = use_of_care,
             y = n,
             fill = marital) ) +
  geom_bar(position = "fill",
           stat = "identity") +
  coord_flip() + scale_fill_manual(values=cp) +
  scale_y_continuous(labels = scales::percent_format())

```

$~$

$~$

$~$

$~$

## Dwelling Location
```{r, message=FALSE}

demo %>% 
  group_by(use_of_care, dwelling) %>% 
  summarise(group_n = n()) %>%
  mutate(percentage = sprintf("%0.1f%%", (group_n/sum(group_n) * 100))) %>% 
  kable(align=rep('c', 4)) %>% 
  kable_styling(bootstrap_options = "striped") %>% 
  pack_rows("Virtual - Never", 1, 3,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - Existing", 4, 6,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - New", 7, 9,
            label_row_css = "background-color: #666; color: #fff;")

demo %>% 
  group_by(use_of_care) %>% 
  count(dwelling) %>% 
  ggplot(aes(x = use_of_care,
             y = n,
             fill = dwelling) ) +
  geom_bar(position = "fill",
           stat = "identity") +
  coord_flip() + scale_fill_manual(values=cp) +
  scale_y_continuous(labels = scales::percent_format())

```

$~$

$~$

$~$

$~$

## Priority Status
```{r, message=FALSE}

demo %>% 
  group_by(use_of_care, priority) %>% 
  summarise(group_n = n()) %>%
  mutate(percentage = sprintf("%0.1f%%", (group_n/sum(group_n) * 100))) %>% 
  kable(align=rep('c', 4)) %>% 
  kable_styling(bootstrap_options = "striped") %>% 
  pack_rows("Virtual - Never", 1, 4,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - Existing", 5, 8,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - New", 9, 12,
            label_row_css = "background-color: #666; color: #fff;")

demo %>% 
  group_by(use_of_care) %>% 
  count(priority) %>% 
  ggplot(aes(x = use_of_care,
             y = n,
             fill = priority) ) +
  geom_bar(position = "fill",
           stat = "identity") +
  coord_flip() + scale_fill_manual(values=cp) +
  scale_y_continuous(labels = scales::percent_format())

```

$~$

$~$

$~$

$~$

## Homeless Status
```{r, message=FALSE}

demo %>% 
  group_by(use_of_care, homeless) %>% 
  summarise(group_n = n()) %>%
  mutate(percentage = sprintf("%0.1f%%", (group_n/sum(group_n) * 100))) %>% 
  kable(align=rep('c', 4)) %>% 
  kable_styling(bootstrap_options = "striped") %>% 
  pack_rows("Virtual - Never", 1, 2,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - Existing", 3, 4,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - New", 5, 6,
            label_row_css = "background-color: #666; color: #fff;")

demo %>% 
  group_by(use_of_care) %>% 
  count(homeless) %>% 
  ggplot(aes(x = use_of_care,
             y = n,
             fill = homeless) ) +
  geom_bar(position = "fill",
           stat = "identity") +
  coord_flip() + scale_fill_manual(values=cp) +
  scale_y_continuous(labels = scales::percent_format())

```

$~$

$~$

$~$

$~$

## Chronic Conditions
```{r, message=FALSE}

demo %>% 
  group_by(use_of_care, chronic) %>% 
  summarise(group_n = n()) %>%
  mutate(percentage = sprintf("%0.1f%%", (group_n/sum(group_n) * 100))) %>% 
  kable(align=rep('c', 4)) %>% 
  kable_styling(bootstrap_options = "striped") %>% 
  pack_rows("Virtual - Never", 1, 3,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - Existing", 4, 6,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - New", 7, 9,
            label_row_css = "background-color: #666; color: #fff;")

demo %>% 
  group_by(use_of_care) %>% 
  count(chronic) %>% 
  ggplot(aes(x = use_of_care,
             y = n,
             fill = chronic) ) +
  geom_bar(position = "fill",
           stat = "identity") +
  coord_flip() + scale_fill_manual(values=cp) +
  scale_y_continuous(labels = scales::percent_format())

```

$~$

$~$

$~$

$~$

## Mental Health Condition
```{r, message=FALSE}

demo %>% 
  group_by(use_of_care, mental) %>% 
  summarise(group_n = n()) %>%
  mutate(percentage = sprintf("%0.1f%%", (group_n/sum(group_n) * 100))) %>% 
  kable(align=rep('c', 4)) %>% 
  kable_styling(bootstrap_options = "striped") %>% 
  pack_rows("Virtual - Never", 1, 2,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - Existing", 3, 4,
            label_row_css = "background-color: #666; color: #fff;") %>% 
  pack_rows("Virtual - New", 5, 6,
            label_row_css = "background-color: #666; color: #fff;")

demo %>% 
  group_by(use_of_care) %>% 
  count(mental) %>% 
  ggplot(aes(x = use_of_care,
             y = n,
             fill = mental) ) +
  geom_bar(position = "fill",
           stat = "identity") +
  coord_flip() + scale_fill_manual(values=cp) +
  scale_y_continuous(labels = scales::percent_format())

```

