---
title: "T-Tests"
author: "Allissa Nguyen"
date: "11/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressPackageStartupMessages(library(tidyverse, quietly = TRUE))
suppressPackageStartupMessages(library(hrbrthemes, quietly = TRUE))

```

```{r}
setwd("~/VA")
vets <- read_csv("csv_files/veterans_table1.csv")
```
```{r}
setwd("~/VA")
visits_mental <- read_csv("csv_files/visits_mental.csv")
```

# T-Tests

Some sample code and explanation/use case information:

https://statsandr.com/blog/student-s-t-test-in-r-and-by-hand-how-to-compare-two-groups-under-different-scenarios/

https://www.r-bloggers.com/2020/03/how-to-do-a-t-test-or-anova-for-many-variables-at-once-in-r-and-communicate-the-results-in-a-better-way/


## Formatting variables
If there is a misclassifcation of object, converts class to "date".

```{r}
visits_mental$visit_day = as.Date(visits_mental$visit_day)
str(visits_mental)
str(vets)
```


## Adding dates in veterans table data
```{r}
visits_mental_joined <- left_join(vets, 
                     select(visits_mental, VID, visit_day), 
                     by = c("VID" = "VID"))
```


## subset data for date range 
```{r}
1 = visits_mental_joined %>%
          filter(between(x = visit_day, 
                         left = as.Date("2020-01-01"), 
                         right = as.Date("2020-01-31")))

2 = visits_mental_joined %>%
          filter(between(x = visit_day, 
                         left = as.Date("2020-05-01"), 
                         right = as.Date("2020-05-31")))
```


# Combine data for selected date range

```{r}
date = rbind(1, 2)
sub = subset(date,
                age == '65 plus')
age$cond = rep(c('before', 'after'), each = 31)
```



### Black veterans before and after

```{r}
library("report")


```




report(test)

### 65+ veterans before and after


## 65+ veterans before and after

### Subset  for 65 plus age

```{r}
sub = subset(date,
                age == '65 plus')
```

### Get counts data for age 65+ 

```{r}
age = sub %>%
          select(age, visit_day) %>%
          group_by(visit_day) %>%
          summarise(count = n()) 

age$cond = rep(c('before', 'after'), each = 31)
head(age)
```

### t-test

```{r}
t.test<-t.test(count ~ cond,
       data = age)
t.test
```


```{r}
report(t.test)
```


```{r}
age %>% 
  ggplot +
  geom_boxplot(mapping = aes(x=cond, y = count)) +
  theme_classic() +
  labs(title = "65+ veterans before and after",
       x = "condition")

```
```{r}
write.csv(age, 'csv')
```


### Ward 7 veterans before and after

```{r}



```







## Ward 7 veterans before and after

### Subset  for ward 7 veterans

```{r}
sub2 = subset(date,
                chronic == '7')
```

### Get counts data for ward 7 veterans 

```{r}
vet7 = sub2 %>%
          select(chronic, visit_day) %>%
          group_by(visit_day) %>%
          summarise(count = n()) 

vet7$cond = rep(c('before', 'after'), each = 31)
head(vet7)
```

### t-test

```{r}
t.test<-t.test(count ~ cond,
       data = vet7)
t.test
```


```{r}
report(t.test)
```

```{r}
write.csv(vet7, 'vet7.csv')
```



### Ward 8 vs. Ward 1 after

```{r}



```


### Ward 8 vs. Ward 1 before

```{r}



```


### White Male 18-34 before vs. Black Male 18-34 before

```{r}



```













# Analysis of variance (ANOVA)

## Jan 2020 vs April 2020, visits by race

### subset data for date range 

```{r}
aov1 = visits_mental_joined %>%
          filter(between(x = visit_day, 
                         left = as.Date("2020-01-01"), 
                         right = as.Date("2020-01-31")))

aov2 = visits_mental_joined %>%
          filter(between(x = visit_day, 
                         left = as.Date("2020-04-01"), 
                         right = as.Date("2020-04-30")))
```

### Combine data for selected date range

```{r}
aov = rbind(aov1, aov2)
```

### Visits by race

```{r}
aov.race = aov %>%
          select(race, visit_day) %>%
          group_by(race, visit_day) %>%
          summarise(count = n())
head(aov.race)
```


### Analysis of variance

```{r}
 = aov(formula = count ~ race, data = aov.race)
anova()
```








<!-- copy of r -->

# Importing data

```{r}
# Importing veterans table data
.veterans = read.csv(file = 'veterans_table1.csv', 
                       header = TRUE)

# Importing visits_mental data
.visits = read.csv(file = 'visits_mental.csv', 
                     header = TRUE)
```

# Formatting variables

```{r}
.visits$visit_day = as.Date(.visits$visit_day)
str(.visits)
```

# Adding dates in veterans table data

```{r}
.comb <- left_join(.veterans, 
                     select(.visits, VID, visit_day), 
                     by = c("VID" = "VID"))
```

# subset data for date range 

```{r}
.comb1 = .comb %>%
          filter(between(x = visit_day, 
                         left = as.Date("2020-01-01"), 
                         right = as.Date("2020-01-31")))

.comb2 = .comb %>%
          filter(between(x = visit_day, 
                         left = as.Date("2020-05-01"), 
                         right = as.Date("2020-05-31")))
```

# Combine data for selected date range

```{r}
.date = rbind(.comb1, .comb2)
```

# T test analysis

## 65+ veterans before and after

### Subset  for 65 plus age

```{r}
.sub = subset(.date,
                age == '65 plus')
```

### Get counts data for age 65+ 

```{r}
.age = .sub %>%
          select(age, visit_day) %>%
          group_by(visit_day) %>%
          summarise(count = n()) 

.age$cond = rep(c('before', 'after'), each = 31)
head(.age)
```

### t-test

```{r}
t.test(count ~ cond,
       data = .age)
```

## Ward 7 veterans before and after

### Subset  for ward 7 veterans

```{r}
.sub2 = subset(.date,
                chronic == '7')
```

### Get counts data for ward 7 veterans 

```{r}
.vet7 = .sub2 %>%
          select(chronic, visit_day) %>%
          group_by(visit_day) %>%
          summarise(count = n()) 

.vet7$cond = rep(c('before', 'after'), each = 31)
head(.vet7)
```

### t-test

```{r}
t.test(count ~ cond,
       data = .vet7)
```

## Ward 8 vs. Ward 1 after

```{r}
.w.after = .date %>%
          filter(between(x = visit_day, 
                         left = as.Date("2020-05-01"), 
                         right = as.Date("2020-05-31")))

.w.after = subset(.w.after, 
                   chronic %in% c('8', '1')) 

.w.after = .w.after %>%
          select(chronic, visit_day) %>%
          group_by(chronic, visit_day) %>%
          summarise(count = n())
head(.w.after)
```

### t-test

```{r}
t.test(count ~ chronic,
       data = .w.after)
```

## White Male 18-34 before vs. Black Male 18-34 before

```{r}
.race.before = .date %>%
          filter(between(x = visit_day, 
                         left = as.Date("2020-01-01"), 
                         right = as.Date("2020-01-31")))

.race = filter(.race.before,
                 sex=='Male' & age =='18-44')

.race2 = filter(.race,
                 race =='White' | race =='Black')


.race3 = .race2 %>%
          select(race, age, sex, visit_day) %>%
          group_by(race, age, sex, visit_day) %>%
          summarise(count = n())
head(.race3)
```

### t-test

```{r}
t.test(count ~ race,
       data = .race3)
```

# Analysis of variance (ANOVA)

## Jan 2020 vs April 2020, visits by race

### subset data for date range 

```{r}
.aov1 = .comb %>%
          filter(between(x = visit_day, 
                         left = as.Date("2020-01-01"), 
                         right = as.Date("2020-01-31")))

.aov2 = .comb %>%
          filter(between(x = visit_day, 
                         left = as.Date("2020-04-01"), 
                         right = as.Date("2020-04-30")))
```

### Combine data for selected date range

```{r}
.aov = rbind(.aov1, .aov2)
```

### Visits by race

```{r}
.aov.race = .aov %>%
          select(race, visit_day) %>%
          group_by(race, visit_day) %>%
          summarise(count = n())
head(.aov.race)
```

### Analysis of variance

```{r}
 = aov(formula = count ~ race, data = .aov.race)
anova()
```

## Ward 8 vs. Ward 1 women by race

### Subsetting data

```{r}
.w.w = filter(.comb,
                 chronic == c('8', '1') & sex =='Female')

.w.w2 = .w.w %>%
          select(chronic, race, sex) %>%
          group_by(chronic, race, sex) %>%
          summarise(count = n())
head(.w.w2)
```

### Analysis of variance

```{r}
2 = aov(formula = count ~ race, data = .w.w2)
anova(2)
```

